## DiagrammeR

you can visit the following sites for details:

* [DiagrammeR-github](https://github.com/rich-iannone/DiagrammeR)
* [DiagrammeR-docs](http://rich-iannone.github.io/DiagrammeR/docs.html)
* [mermaid-docs](https://mermaidjs.github.io/)


```{r setup, include = FALSE}
    library(DiagrammeR)
    library(dplyr)
```

### flowchart

node directions:

* TB,TD - top bottom
* BT - bottom top
* RL - right left
* LR - left right

node shapes:

* id[start now] - rectangle
* id(start now) - rectangle with round edges
* id((start now)) - circle
* id>start now] - asymmetric
* id{start now} - rhombus

node links:

* `-->` - arrow head
* `---` - open link
* `a-- this is the text ---b` or `a---|this is the text|b`- text on open links
* `a-- this is the text -->b` or `a-->|this is the text|b`- text on arrow head links
* `a-.->b` - dotted
* `a-. text .->` - dotted with text
* `a ==> b` - thick link
* `a == text ==> b` - thick link with text

subgraphs:

    subgraph title
        graph definition
    end

```{r example}
    DiagrammeR("graph TB
                    id1[start now] -- arrow head --> rectangle
                    id2(start now) -- open link --- round
                    id3((start now)) -. dotted .-> circle
                    id4>start now] == thick link ==> asymmetric
                    id5{start now} --> rhombus
               ")

```

### graphviz

graph:

* layout:
    * dot
    * neato
    * twopi
    * circo
* rankdir:
    * LR
    * TB
* overlap:
    * true
    * false

node attributes:


```{r}
grViz("
      digraph node_attributes {

        graph [rankdir = LR]
        
        # 
        node [color=red shape=box]

        root [label = 'node']
        
        node []
        shape
        
        # node shape
        node []
        shape_oval; shape_box; shape_square; shape_circle
        shape_egg; shape_triangle; shape_diamond; shape_polygon;
        shape_point; shape_plaintext

        shape_oval[shape=oval, label='oval']
        shape_box[shape=box, label='box']
        shape_square[shape=square, label='square']
        shape_circle[shape=circle, label='circle']
        shape_egg[shape=egg, label='egg']
        shape_triangle[shape=triangle, label='triangle']
        shape_diamond[shape=diamond, label='diamond']
        shape_polygon[shape=polygon, label='polygon']
        shape_point[shape=point, label='point']
        shape_plaintext[shape=plaintext, label='plaintext']

        edge [color=gray]

        root -> shape -> {shape_oval, shape_box, shape_square, shape_circle, shape_egg; shape_triangle, shape_diamond, shape_polygon, shape_point, shape_plaintext}
      }
      ")
```


* color - provide an X11 or hexadecimal color (append 2 digits to hex for alpha)
* style - usually given the value filled if you'd like to fill a node with a colors
    * filled
* fillcolor - when `style=filled` is set
* fontcolor
* alpha
* fontname
* fontsize
* shape
    * oval - default
    * square
    * box
    * circle
    * diamond
    * egg
    * plaintext
    * point
    * triangle
    * polygon
* penwidth
* sides - if `shape = polygon`, the number of sides can be provided here
* distortion - the node distortion for any `shape = polygon`
* peripheries - the number of peripheries (essentially, additional shape outlines)
* fixedsize
    * true
    * false
* x - the x position of the node (requires graph attr layout = neato to use)
* y - the y position of the node (requires graph attr layout = neato to use)
* height
* width
* tooltip
* icon

edge attributes:

* arrowsize
* arrowhead
    * normal - default
    * box
    * crow
    * curve
    * diamond
    * dot
    * inv
    * none
    * tee
    * vee
* arrowtail
* color
* dir - direction
    * forward
    * back
    * both
    * none
* alpha
* headport
* tailport
* fontname
* fontsize
* fontcolor
* label - label text for the line between nodes
* penwidth
* minlen - minimum rank distance between head and tail
* tooltip

```{r}
    grViz("
          digraph boxes_and_circles {
            # a 'graph' statement
            graph [overlap = true, fontsize = 10]
            
            node [shape = box,
                  fontname = Helvetica]
            A; B; C; D; E; F
            
            node [shape = circle,
                  fixedsize = true,
                  width = 0.9]
            1; 2; 3; 4; 5; 6; 7; 8

            A->1 B->2 B->3 B->4 C->A
            1->D E->A 2->4 1->5 1->F
            E->6 4->6 5->7 6->7 3->8
          }
          ")
```
