## DiagrammeR

you can visit the following sites for details:

* [DiagrammeR-github](https://github.com/rich-iannone/DiagrammeR)
* [DiagrammeR-docs](http://rich-iannone.github.io/DiagrammeR/docs.html)
* [mermaid-docs](https://mermaidjs.github.io/)


```{r setup, include = FALSE}
    library(DiagrammeR)
    library(dplyr)
```

### flowchart

node directions:

* TB,TD - top bottome
* BT - bottom top
* RL - right left
* LR - left right

node shapes:

* id[start now] - rectangle
* id(start now) - with round edges
* id((start now)) - circle
* id>start now] - asymmetric
* id{start now} - rhombus

node links:

* `-->` - arrow head
* `---` - open link
* `a-- this is the text ---b` or `a---|this is the text|b`- text on open links
* `a-- this is the text -->b` or `a-->|this is the text|b`- text on arrow head links
* `a-.->b` - dotted
* `a-. text .->` - dotted with text
* `a ==> b` - thick link
* `a == text ==> b` - thick link with text

subgraphs:

    subgraph title
        graph definition
    end




```{r example}
    DiagrammeR("graph TB
                    id1[start now] -- arrow head --> rectangle
                    id2(start now) -- open link --- round
                    id3((start now)) -. dotted .-> circle
                    id4>start now] == thick link ==> asymmetric
                    id5{start now} --> rhombus
               ")

```
